---
title: "narG_database"
author: "Benjamin D. Peterson"
date: "2023-05-18"
output: html_document
---

```{r setup, include = FALSE, warning = FALSE}
rm(list = ls())
library(Biostrings)
library(DECIPHER)
library(ggtree)
library(stringr)
library(tidyverse)
library(treeio)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(echo = TRUE)
```

# Generate metadata

First I combined the additional metadata files with the original one, which I saved out.

```{r narG_metadata}
narg_metadata <- read.table("NarG_luke_database_metadata_original.tsv",
                            sep = '\t',
                            header = TRUE)
narg_metadata_add <- read.table("NarG_luke_database_metadata_additional.tsv",
                                sep = '\t',
                                header = TRUE)
narg_metadata_all <- full_join(narg_metadata, narg_metadata_add)
rm(narg_metadata, narg_metadata_add)
write.table(x = narg_metadata_all,
            file = "NarG_luke_database_metadata.tsv",
            row.names = FALSE,
            quote = FALSE)
```

I then generated a vector to replace the tip labels in the tree.

```{r tip_labels}
tip_labels_to_use <- paste(narg_metadata_all$type, ": ", narg_metadata_all$organism,
                           sep = "")
names(tip_labels_to_use) <- narg_metadata_all$geneID
tip_labels_to_use <- tip_labels_to_use[which(tip_labels_to_use != "NA: NA")]
```

Finally, I read in the tree and replaced some of the labels.
I midpoint-rooted the tree as a first pass.

```{r visualize_narG_tree}
unrooted_tree <- read.newick("NarG_luke_database.tree")
add_metadata_index <- which(unrooted_tree$tip.label %in% names(tip_labels_to_use))

unrooted_tree$tip.label[add_metadata_index] <- tip_labels_to_use[unrooted_tree$tip.label[add_metadata_index]]
rooted_tree <- phangorn::midpoint(unrooted_tree)

narg_tree <- ggtree(rooted_tree) +
  geom_tiplab(size=2.5) +
  geom_treescale(x = 5,
                 y = 1)
narg_tree +
  geom_text2(aes(subset = !isTip,
                 label = node),
             size = 2.5)
```

The labeled tips are the *nxrA* sequences, while the cluster with no additional metadata represents the *narG* sequences.
Here's the tree again with just *narG* sequences shown (*nxrA* collapsed).

```{r narG_tree}
collapse(narg_tree, 55, 'mixed') %>%
  collapse(50, 'mixed') %>%
  collapse(53, 'mixed')
```

We'll use this data to generate our *narG* trees to make sure we're not picking up *nxrA* sequences.
